\documentclass[]{scrartcl}

\usepackage{\string~"/LaTeX/StylePackage"}

\title{Lecture 1}
\author{}
\date{\today}


\begin{document}

\maketitle
\newpage
\tableofcontents
\newpage

\subsection{Thermodynamics and Statistical Physics differences}
Thermodynamics:
\begin{itemize}
	\item Macroscopic
	\item Continuous Matter
	\item differentiable
	\item Necessary relations based on some axioms
	\item All properties of matter ($\Delta H_m, \Delta S_v, c_v, \lambda, D$) must be measured
\end{itemize}
Statistical Physics
\begin{itemize}
	\item microscopic
	\item discrete particles
	\item Mechanics
	\item statistical behavior of simplified models
	\item Bottom up explanation of thermodynamics
	\item Properties of model matter ($\Delta H_m, \Delta S_v, c_v, \lambda, D$) can be calculated or measured in a simulation
\end{itemize}

\subsection{Thermodynamics is omnipresent}
\begin{itemize}
	\item Climate: basis for modelling
	\item Sustainability:
	\begin{itemize}
		\item Efficiency in the process industry
		\item Energy efficient heating and cooling of buildings
		\item Efficient engines and refrigerators
	\end{itemize}
	\item Basic research:
	\begin{itemize}
		\item Condensed matter physics
		\item High energy physics
		\item Cosmology
		\item Biophysics
		\item Meteorology
		\item Geology
		\item Chemistry
	\end{itemize}
\item $\Rightarrow$ Relevant for jobs
\end{itemize}
\subsection{Statistical Mechanics, the secret weapon of physicists}
\begin{itemize}
	\item Make microscopic models of new phenomena
	\item Models follow laws of physics and statistics
	\item Derive macroscopic models
\end{itemize}

\section{Basic Concepts}

$T$ is the temperature of an object.\\
If you have an object with $T_1$ and one with $T_2$, then they are at thermal equilibrium iff $T_1 = T_2$.

Heat is related to Energy, such that heat is the microscopic kinetic energy. In a macroscopic picture, heat is the transport of Energy. Heat is written as $Q$.

Work is the mechanical energy acting on a system. It is written as $W$. Mechanically, it is $W = F\Delta x$. Work is an Energy.

A system is isolated from the rest of the world, but there can be energy exchange between them. They can exchange work or heat for example. The internal energy of a System is denoted as $U, E$.

\subsection{First Law of Thermodynamics}
Energy is conserved, that is, energy cannot be destroyed or created, only exchanged.

Any change of energy in a system is caused by a transfer of energy with the outside world. That being, $\Delta U = Q + W$.


\end{document}

